A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-05-18 meth.tex
dh fig and analysis added

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory. In most cases, however, models do report values in the
harmonization year. When model and historical values are relatively close, a
convergence method is provided in order to be as representative as possible to
the underlying unharmonized model results. The final model-based characteristic
of import is whether the model reports negative emissions. Such a case is
possible for gas species which can be extracted from the enviornment, as is the
case for CO2 in future scenarios with high mitigation strategies.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{cov.pdf}
    \caption[]{
      \label{fig:cov}
      The distribution of $\Omega$ values for LUC and non-LUC historical
      trajectories are shown. CEDS historical data is used for non-LUC data and
      \TODO{cite LUC data} is used for LUC data. All historical data has been
      aggregated from countries to IAM model regional definitions, and all gas
      species included in the historical datasets are included in the
      analysis. The solid black line indicates the threshold value used by
      default in \code{aneris}.  
    }
  \end{center}
\end{figure}

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization year.

>>>>

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory. In most cases, however, models do report values in the
harmonization year. When model and historical values are relatively close, a
convergence method is provided in order to be as representative as possible to
the underlying unharmonized model results. The final model-based characteristic
of import is whether the model reports negative emissions. Such a case is
possible for gas species which can be extracted from the enviornment, as is the
case for CO2 in future scenarios with high mitigation strategies. If a model
provides a trajectory that transitions from positive to negative emissions, then
a convergence method is used in order to guarantee capture of this
transition. Otherwise, it is possible, given the relative values between history
and model in the harmonization year, for a negative trajectory to be
inappropriatley harmonized to positive, but decreasing, trajectory.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{cov.pdf}
    \caption[]{
      \label{fig:cov}
      The distribution of $\Omega$ values for LUC and non-LUC historical
      trajectories is shown. CEDS historical data is used for non-LUC data and
      \TODO{cite LUC data} is used for LUC data. All historical data has been
      aggregated from countries to IAM model regional definitions, and all gas
      species included in the historical datasets are included in the
      analysis. The solid black line indicates the threshold value used by
      default in \code{aneris}.  
    }
  \end{center}
\end{figure}

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization year. In order to investigate
the possible values that these relative differences can take, the flagship model
values used in the SSP \TODO{CITE} and (ongoing) CMIP6 intercomparison exercises
are used. A distribution of these differences for all models in the study is
presented in Figure \ref{fig:dh}. Given the available data, a threshold value of
50\% was chosen to be used as a default in \code{aneris}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{dh.pdf}
    \caption[]{
      \label{fig:dh}
      The distribution of relative differences between model and historical
      values in the harmonization year is shown. The solid black line indicates
      the threshold value used by default in \code{aneris}.  
    }
  \end{center}
\end{figure}
